---
title: "Homework 2"
author: "Sunrise Gao"
date: "4/27/2022"
output: html_document
---

```{r setup, echo=FALSE}
library(knitr)
library(MASS)
library(ggplot2)
library(ggformula)
library(tidyverse)
# set global chunk options: images will be 7x5 inches
knitr::opts_chunk$set(fig.width=7, fig.height=5)
options(digits = 4)


## indents are for indenting r code as formatted text
## They may need to be adjusted depending on your OS
# if your output looks odd, increase or decrease indent
indent1 = '    '
indent2 = '        '
indent3 = '            '
```

## 2
(c)
```{r}

x<-c("15","15","15","15","15",
     "20","20","20","20","20",
     "25","25","25","25","25",
     "30","30","30","30","30",
     "35","35","35","35","35")
y<-c(7, 7, 15, 11,9,
     12,17,12,18,18,
     14,19,19,18,18,
     19,25,22,19,23,
     7, 10,11,15,11)

cotton_data <- data.frame(x, y)

cotton_data

aov <- aov(y~x, cotton_data)
summary(aov)

```



```{r}

res = aov$residuals

qqnorm(res)
qqline(res)

plot(res, ylab="Error")
abline(h=0, col="red")

plot(aov, which = 1)




```
Plot 1: Points lie and are fairly close to the line - Normality assumption is met.
Plot 2: There is no pattern shown as points spread randomly - Independence assumption is met.
Plot 3: a. points are symmetric; b. points spread randomly (no pattern); c. the red line is horizontal - Constant variance assumption is met and the model is adequate.

```{r}
x<-c("1","1","1",
     "2","2","2",
     "3","3","3",
     "4","4","4")
y<-c(2.76,5.67,4.49,
     1.43,1.70,2.19,
     2.34,1.97,1.47,
     0.94,1.36,1.65)

wafer_data <- data.frame(x, y)
aov2 <- aov(y~x, wafer_data)
summary(aov2)

```



```{r}

res2 = aov2$residuals

qqnorm(res2)
qqline(res2)

plot(res2, ylab="Error")
abline(h=0, col="red")

plot(aov2, which = 1)




```
Plot 1: Points lie and are fairly close to the line - Normality assumption is met.
Plot 2: There is a wave-like pattern shown as points spread - Independence assumption is not met.
Plot 3: a. points are symmetric; b. points spread in a non-constant variance band; c. the red line is not horizontal, there is an obvious linear trend - Constant variance assumption is not met and the model is not adequate.
